---
title: "Substandard Deviations Analysis"
author: "Javier Valcarcel & Justin Shalap"
date: "September 20, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Wrangling

## Load relevant R libraries

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
library(ggplot2)
library(hrbrthemes)
library(gcookbook)
```

```{r}
sales <- read_excel("product_sales.xlsx", sheet = 1)

head(sales)
```

## Data transformation of Voltswitch sales

We see the structure of the data and wish to transform it into a format that can be merged with
other licensee sales data for comparison and regression. Specifically we want to know their total
sales, total sales of products using InfiniCharge and what percentage of their revenue came from
infringing technology.



```{r, warning = FALSE}
mycro_line <- c("Mycro 1", "Mycro 2", "Mycro 3")

total_sales_volt <- sales %>%
  filter(company == "Voltswitch") %>%
  mutate(year = year(month)) %>%
  group_by(year) %>%
  summarise(total_sales = sum(total_sales_USD))

total_sales_volt_mycro <- sales %>%
  filter(product %in% mycro_line) %>%
  mutate(year = year(month)) %>%
  group_by(year) %>%
  summarise(total_sales_mycro  = sum(total_sales_USD))

volt_sales <- merge(total_sales_volt, total_sales_volt_mycro)

volt_sales1 <- volt_sales %>%
  mutate(percent_patent = total_sales_mycro / total_sales)

head(volt_sales1)
```


## Data transformation of licensee sales

From here we load, clean and merge the sales data from the other licensees to merge with the
Voltswitch data.



```{r}
companies <- c("Year", "Rekall", "Tyrell Corp.", "Sirius Cybernetics",
               "Cyberdyne Systems", "Nakatomi Corp.")

licensed_sales <- read_excel("sales_data.xlsx", sheet = 1, cell_rows(7:14))

licensed_sales_clean <- licensed_sales[, c(1, 3, 4, 5, 6, 7)]
names(licensed_sales_clean) <- companies

licensed_patent_clean <- licensed_sales[, c(1, 9, 10, 11, 12, 13)]
names(licensed_patent_clean) <- companies

lsc <- licensed_sales_clean %>%
  gather(company, sales, -Year)

lpc <- licensed_patent_clean %>% 
  gather(company, percent, -Year)

clean_sales_patents <- merge(lsc, lpc) %>%
  arrange(Year, company)

head(clean_sales_patents)
```


## Merge Voltswitch with other licensees

We now are able to merge the Voltswitch data with other licensees to view a
combined data set ready for regression.



```{r}
volt_sales2 <- volt_sales1 %>%
  mutate(Year = year) %>%
  mutate(company = "Voltswitch") %>%
  mutate(sales = total_sales) %>%
  mutate(percent = percent_patent) %>%
  select(Year, company, sales, percent)

total_sales_df <- rbind(clean_sales_patents, volt_sales2) %>% arrange(Year, company)

royalty_schedule <- tibble(
  company = companies[-1],
  royalty_percent = c(0.05, 0.07, 0.02, 0.03, 0.01))

reg_tot1 <- total_sales_df %>%
  merge(royalty_schedule) %>%
  mutate(royalties_due = sales * percent * royalty_percent)

reg_tot2 <- reg_tot1 %>%
  group_by(company) %>%
  summarise(tot_sales = sum(sales),
            avg_percent = mean(percent),
            royalty_percent = mean(royalty_percent))

head(reg_tot2)
```


# Analysis and Modeling

## Estimate reasonable royalties

Here we use linear regression to estimate the best fit given the sales data and 
share of sales by InfiniCharge products. We also calculate the royalties due.



```{r}
roy_mod <- lm(royalty_percent ~ tot_sales + avg_percent, reg_tot2)

volt_sales3 <- volt_sales2 %>%
  summarise(company = "Voltswitch",
            tot_sales = sum(sales),
            avg_percent = mean(percent))

pred_roy_percent <- predict(roy_mod, volt_sales3)

volt_sales3$royalty_percent <- pred_roy_percent

reg_tot3 <- rbind(reg_tot2, volt_sales3)

reg_tot3 %>%
  mutate(royalties_due = tot_sales * avg_percent * royalty_percent)


```

## Validation of reasonable royalties to defeat Georgia-Pacific defense.

We look to validate that the estimated royalty given the presence of similar licenses
is profitable to demonstrate mutual gains to trade for the license. We should expect the price
to be above the cost of the license since there are no close substitutes. We use the split-
apply-combine methodology.

```{r}

results <- read_excel("survey_results.xlsx", sheet = "Survey Results", cell_rows(4:54))

key <- read_excel("survey_results.xlsx", sheet = "Question Key", cell_rows((4:28)))

results1 <- results[, -1]

split_prices <- split(results1, seq(1:50))
prices_lists <- lapply(split_prices, t)
prices_lists <- lapply(prices_lists, cbind, key)
tester <- bind_rows(prices_lists)

tester_names <- c("Price", "Question", "Talk_Time", "Weight", "PPI", "Camera")
names(tester) <- tester_names

head(tester)

tester_model <- lm(Price ~ Talk_Time + Weight + PPI + Camera, tester)

tester_model

anova(tester_model)
```



## Predicting prices for different products

Given the model for prices based on the survey results, we now use that to price
best alternatives to satisfy the but-for process. But-for the infringement
what would Voltswitch's alternatives have been? We price those and analyze whether
those alternatives were as profitable as using InfiniCharge.


```{r}
desc <- read_excel("product_descriptions.xlsx", sheet =1, cell_rows(6:23))
marg <- read_excel("product_margins.xlsx", sheet =1, cell_rows(6:23))

desc1 <- desc[, -c(2,5)]
marg1 <- marg[, -2]

desc_marg <- merge(desc1, marg1)
names(desc_marg) <- c("Product", "Parent_Co", "Release_Date", "Talk_Time", "Weight",
                      "Screen", "Memory", "Camera", "PPI", "Margin")

head(desc_marg)

dm_data <- desc_marg %>%
  select(Talk_Time, Weight, PPI, Camera)

prod_preds <- predict(tester_model, dm_data)
desc_marg$pred <- prod_preds

desc_marg <- desc_marg %>%
  mutate(profits = pred * Margin) %>%
  mutate(pred_royalty = pred * 0.02) %>%
  mutate(profit_over_license = profits - pred_royalty) %>%
  arrange(profits)

tail(desc_marg, n= 9)
```

# Validation through visualization

We create some visuals which convey the structure of the data to
non-technical stakeholders

```{r, warning=FALSE}

reg_tot3 %>%
  ggplot(aes(x = company, y = avg_percent, size = royalty_percent)) +
    guides(size = FALSE, colour = FALSE) +
    labs(x = "Companies",
         y = "Yearly average percentage",
         title = "Companies by average share",
         subtitle = "Yearly average from 2009-2014",
         caption = "We see that generally the royalty rate increases as the percentage increases.") +
    geom_point(aes(colour = company)) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 15, hjust = 0.7, size = 10))

```

```{r}
prof_dev <- desc_marg[9:17,]

pd_viz_df <- prof_dev[, c(1, 12)]

pd_viz_df %>%
  mutate(Product = as.factor(Product)) %>%
  ggplot(aes(x = Product, y = profits)) +
    guides(size = FALSE, colour = FALSE) +
    labs(x = "Product",
         y = "Profits per device",
         title = "Profit comparison by products",
         subtitle = "InfiniCharge devices and alternatives",
         caption = "We see that the profits of using InfiniCharge are greater per device than alternative") +
    geom_point(aes(colour = Product, size = profits)) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 15, hjust = 0.7, size = 10))
```